Designing Modular Software: A Case Study in Introductory Statistics

### 2016-10-10 jcgs-paper-2016/paper.Rmd
Typo fixes

Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation and helps abstract away some of the more challenging programming aspects of working directly with R. Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application. Several recent software packages have built up Shiny to provide a web-based system based on R. One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge. Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting. Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis. While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student. An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}. Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default. For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.

@@

The modules included in `intRo` are split into three higher level categories - data, summaries, and inference. Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface. More modules can easily be added by an instructor, as detailed in section \ref{modules}. The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

@@

Shiny observers are a class of reactive objects within the Shiny paradigm which do not return a value [@shiny]. For further discussion of reactivity, see section \ref{reactive-design-choices}. In this example, observers are created to ensure that the choices of variable for the `nonparametric` module are only numeric variables. This is accomplished by utilizing the global reactive `intro.numericnames()`, which returns a character vector containing the variables in the current dataset that are numeric. Finally, there is an event observer to store code generated from the module into the overall code script upon clicking the store button. The presence of this observer code and the definition of the button in the user interface are enforced, and must be present in any `intRo` module.

@@

Reproducibility in `intRo` is accomplished with the previously mentioned `interpolate` function. `interpolate` accepts an expression and an arbitrary number of arguments as an argument, substitutes the arguments into the expression, prints the results to the console, and evaluates the parsed expression. This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`. One potential drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).

>>>>

Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation and helps abstract away some of the more challenging programming aspects of working directly with R. Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application. Several recent software packages have built upon Shiny to provide a web-based system based on R. One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge. Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting. Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis. While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student. An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}. Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default. For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.

@@

The modules included in `intRo` are split into three higher level categories - data, summaries, and inference. Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface. More modules can easily be added by an instructor, as detailed in section \ref{designing-for-modularity}. The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

@@

Shiny observers are a class of reactive objects within the Shiny paradigm which do not return a value [@shiny]. For further discussion of reactivity, see section \ref{designing-for-reactivity}. In this example, observers are created to ensure that the choices of variable for the `nonparametric` module are only numeric variables. This is accomplished by utilizing the global reactive `intro.numericnames()`, which returns a character vector containing the variables in the current dataset that are numeric. Finally, there is an event observer to store code generated from the module into the overall code script upon clicking the store button. The presence of this observer code and the definition of the button in the user interface are enforced, and must be present in any `intRo` module.

@@

Reproducibility in `intRo` is accomplished with the previously mentioned `interpolate` function. `interpolate` accepts an expression and an arbitrary number of arguments as an argument, substitutes the arguments into the expression, prints the results to the console, and evaluates the parsed expression. This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`. A possible drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).
